Density-dependent selection in evolutionary genetics: a lottery model of Grime's triangle

### 2017-01-05 manuscript.tex
latest

As we will see, a generalized version of the classic lottery model of \cite{chesson_1981} is well suited for this purpose. The lottery model gets its name from the way in which it represents competition: the growth of each type depends on how many territories it secures from a fixed territorial stock, which is determined by the number of contestants that it produces relative to the other types in the population weighted by a coefficient for each type representing competitive ability. This is much simpler than having coefficients for the pairwise effects of types on each other (e.g. the $\alpha$ coefficients in the generalized Lotka-Volterra equations), or than modeling resource consumption explicitly \citep{tilman_1982}. The classic lottery model  breaks down at low absolute abundances at the scale of territorial contests (section ``Mean field approximation''), but this was not important for its original application to reef fishes, where a huge number of larvae from each type compete to secure territories each generation \citep{chesson_1981}. Here we analytically extend the classic lottery model to correctly account for low density behavior.

>>>>

As we will see, a generalized version of the classic lottery model of \cite{chesson_1981} is well suited for our purposes. In the classic lottery model, mature individuals (henceforth ``adults'') each require their own territory, whereas newborn individuals (henceforth ``propagules'') disperse to, and subsequently compete for, the territories made available by the death of adults. Territorial contest among propagules leaves a single victorious adult per territory, with the victor chosen at random from the propagules present (akin to a lottery), with probabilities weighted by a coefficient for each type representing competitive ability. This representation of competition is much simpler than having coefficients for the pairwise effects of types on each other (e.g. the $\alpha$ coefficients in the generalized Lotka-Volterra equations), or than modeling resource consumption explicitly \citep{tilman_1982}. The classic lottery model also has similarities with the Wright-Fisher model in population-genetics (section BLAH).

However, the classic lottery model breaks down at low densitites, specifically if there are only  a few propagules dispersing to each territory (section ``Mean field approximation''). This was not a limitation in its original application to reef fishes, where a huge number of larvae from each species compete to secure territories each generation \cite{chesson_1981}, but is a critical limitation here. We analytically extend the classic lottery model to correctly account for low density behavior.

%(each generation is a sample of constant size, with replacement, from the previous one)
